How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-23 research/paper/draft-editing.tex

This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.

@@

@@

One major concern about photo-$z$s is the intrinsic scatter of point estimators, including those derived from photo-$z$ PDFs, that is observed to varying extents with every existing photo-$z$ algorithm and illustrated in Fig. \ref{fig:intscat}.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying intrinsic scatter]}
		\label{fig:intscat}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due to the dearth of distant galaxies with spectroscopic redshifts.  The interim prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of training-like (low-$z$ favoring) interim prior]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  Here we show that the traditional estimators of $n(z)$ cannot recover a highly featured $n(z)$.  The implication of this failure is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.

In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has nontrivial structure.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of a highly featured (delta function or multimodal?) true $n(z)$]}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}

>>>>

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous function with known parameters $\vec{\theta}$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of $N$ true redshifts $z_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} | z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters $\vec{\phi}^{*}$ of the interim prior.
	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | \vec{\phi}^{*})$ in this parametrization.
\end{enumerate}

In all of the following validation tests, we use $10$ bins and $N = 10,000$ galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This method for deriving mock data is referred to as the fiducial case, and variations on it will refer directly to the steps that are altered.

@@

Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  Here we show that the traditional estimators perform better when the true $n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The implication of this issue is quite serious; the consistently smooth, unimodal $n(z)$ estimates appearing in the literature could result from much more featured true redshift distributions, and there would be no way to catch this error without using a fully probabilistic method.

\subsubsection{Featureless $n(z)$}
\label{sec:smooth}

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 photo-$z$ PDFs.  [check and include citation!]

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
		\caption{All estimators perform well when the true $n(z)$ is well behaved, exhibiting significant deviation only when $n(z)$ is very small, as the sample size of true redshifts in that range will be small.  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with nontrivial structure.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has been quoted as what LSST, etc. needs, but that's not what comes out of photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$

@@

One major concern about photo-$z$s is the intrinsic scatter of point estimators, including those derived from photo-$z$ PDFs, that is observed to varying extents with every existing photo-$z$ algorithm and illustrated in Fig. \ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case to simply broaden the single Gaussian component of the likelihood.  To enforce self-consistency, the mean is a random variable drawn from a Gaussian distribution with the newly increased variance.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
		\caption{As the intrinsic scatter increases, the discrepancy between estimators increases.  In particular, the stacked estimator and marginalized point estimators predict $\ln[n(z)]$ to be smoother than the truth, while the  [This would be a lot more compelling with more galaxies.  Also, the weird edge effects in the top panel are real, because the point estimator is the MAP, not the center of the Gaussian likelihood, and there's no requirement that the mean of the likelihood be within the true redshift range.]  Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in solid lines, with a dotted line indicating zero probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a continuous function and under a binned parametrization.}
		\label{fig:intscat}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due to the dearth of distant galaxies with spectroscopic redshifts.  The interim prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior corresponding to a training set biased in this way.  We chose an interpolation of the interim prior used for the SDSS DR8 photo-$z$ PDFs.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\
		\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
		\caption{[The case of a low-$z$ favoring interim prior was a very compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
% 		\includegraphics[width=0.5\textwidth]{money.png}
		\caption{[moneyplot of error ellipses resulting from different $n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}

### 2017-06-23 research/paper/draft.tex

@@

One major concern about photo-$z$s is the intrinsic scatter of point 
estimators, including those derived from photo-$z$ PDFs, that is observed to 
varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
\ref{fig:intscat}.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with varying 
intrinsic scatter]}
		\label{fig:intscat}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with template-like 
catastrophic outliers]}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of $z_{phot}$ vs. $z_{spec}$ with training-like 
catastrophic outliers]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of template-like (multimodal) interim prior]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due 
to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
corresponding to a training set biased in this way.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of training-like (low-$z$ favoring) interim 
prior]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
Here we show that the traditional estimators of $n(z)$ cannot recover a highly 
featured $n(z)$.  The implication of this failure is quite serious; the 
consistently smooth, unimodal $n(z)$ estimates appearing in the literature 
could result from much more featured true redshift distributions, and there 
would be no way to catch this error without using a fully probabilistic method.

In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured}, which has 
nontrivial structure.

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[plot of a highly featured (delta function or 
multimodal?) true $n(z)$]}
		\label{fig:featured}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		%\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{[moneyplot of error ellipses resulting from different 
$n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}

>>>>

Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
Here we show that the traditional estimators perform better when the true 
$n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by 
experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The 
implication of this issue is quite serious; the consistently smooth, unimodal 
$n(z)$ estimates appearing in the literature could result from much more 
featured true redshift distributions, and there would be no way to catch this 
error without using a fully probabilistic method.

\subsubsection{Featureless $n(z)$}
\label{sec:smooth}

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
photo-$z$ PDFs.  [check and include citation!]

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
		\caption{All estimators perform well when the true $n(z)$ is 
well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
as the sample size of true redshifts in that range will be small.  Top panel: 
The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
lines, with a dotted line indicating zero probability.  Bottom panel: Various 
estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with 
nontrivial structure.

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the 
intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
been quoted as what LSST, etc. needs, but that's not what comes out of 
photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
solid lines, with a dotted line indicating zero probability.  Bottom panel: 
Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$

@@

One major concern about photo-$z$s is the intrinsic scatter of point 
estimators, including those derived from photo-$z$ PDFs, that is observed to 
varying extents with every existing photo-$z$ algorithm and illustrated in Fig. 
\ref{fig:intscat}.  To emulate intrinsic scatter, we modify the fiducial case 
to simply broaden the single Gaussian component of the likelihood.  To enforce 
self-consistency, the mean is a random variable drawn from a Gaussian 
distribution with the newly increased variance.

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/high_scatter/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/high_scatter/estimators.png}		
		\caption{As the intrinsic scatter increases, the discrepancy 
between estimators increases.  In particular, the stacked estimator and 
marginalized point estimators predict $\ln[n(z)]$ to be smoother than the 
truth, while the  [This would be a lot more compelling with more galaxies.  
Also, the weird edge effects in the top panel are real, because the point 
estimator is the MAP, not the center of the Gaussian likelihood, and there's no 
requirement that the mean of the likelihood be within the true redshift range.] 
 Top panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
solid lines, with a dotted line indicating zero probability.  Bottom panel: 
Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
as a continuous function and under a binned parametrization.}
		\label{fig:intscat}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/template_outliers/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/template_outliers/estimators.png}

		\caption{Top panel: The traditional MAP reduction of a 
photo-$z$ PDF against the true redshifts with a few rescaled photo-$z$ interim 
posteriors are overplotted in solid lines, with a dotted line indicating zero 
probability.  Bottom panel: Various estimators of $\ln[n(z)]$, the interim 
prior, and the true $\ln[n(z)]$ as a continuous function and under a binned 
parametrization.}
		\label{fig:tempcatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/training_outliers/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/training_outliers/estimators.png}	
		\caption{[This has proven to be the most challenging test case 
to implement, and there's clearly still a bug in the function that makes the 
likelihoods.]}
		\label{fig:traincatout}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/template_prior/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/template_prior/estimators.png}	
		\caption{[The case of a multimodal interim prior was a very 
compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:tempintpr}
	\end{center}
\end{figure}

@@

An interim prior based on a training set may be biased toward low redshifts due 
to the dearth of distant galaxies with spectroscopic redshifts.  The interim 
prior shown in Fig. \ref{fig:trainintpr} is an emulation of an interim prior 
corresponding to a training set biased in this way.  We chose an interpolation 
of the interim prior used for the SDSS DR8 photo-$z$ PDFs.

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/training_prior/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/training_prior/estimators.png}	
		\caption{[The case of a low-$z$ favoring interim prior was a 
very compelling test in the previous version but somehow isn't anymore.]}
		\label{fig:trainintpr}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\caption{[moneyplot of error ellipses resulting from different 
$n(z)$ estimators]}
		\label{fig:money}
	\end{center}
\end{figure}
